---
title: "MA_Hw_04_01"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---
## Packages
```{r}
suppressMessages(library(plyr))
suppressMessages(library(rethinking))
suppressMessages(library(rstan))
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
```

## Data-processing
### Read data
```{r}
data("rugged")
rugged
```

### Full_df & Without_Sey_df
Form two dataframe, one for all data, and another for data without Seychelles
```{r}
full_data <- rugged %>%
  dplyr::select(country, rgdppc_2000, cont_africa, rugged) %>% 
  mutate(AR = cont_africa*rugged)
full_data <- full_data[complete.cases(full_data),] #complete.cases for selecting non-NA row
data_without_sey <- full_data %>% filter(country != "Seychelles")
data_without_sey
```

### Standardlization
```{r}
full_data = full_data %>% 
  mutate(
    log_gdp = log(rgdppc_2000)) %>% 
  mutate(
    log_gdp_std = log_gdp / mean(log_gdp),
    rugged_std = rugged / max(rugged))

data_without_sey = data_without_sey %>% 
  mutate(
    log_gdp = log(rgdppc_2000)) %>% 
  mutate(
    log_gdp_std = log_gdp / mean(log_gdp),
    rugged_std = rugged / max(rugged))

full_data
data_without_sey
```

## Part A
### Model
```{r}
q1_m1 <- "
data{
  int N;
  vector[N] y;
  vector[N] A;
  vector[N] R;
  vector[N] AR;
}

parameters{
  real alpha;
  real beta_a;
  real beta_r;
  real beta_ar;
  real sigma;
}

model{
  vector[N] mu = alpha + beta_a*A + beta_r*R +beta_ar*AR;
  y ~ normal(mu, sigma);

  alpha ~ normal(1, 0.1);
  beta_a ~ normal(0, 0.3);
  beta_r ~ normal(0, 0.3);
  beta_ar ~ normal(0, 0.3);
  sigma ~ exponential(1);
}

generated quantities{
  vector[N] pred_mu;
  real pred_y[N];
  vector[N] log_lik;

  pred_mu = alpha + beta_a*A + beta_r*R + beta_ar*AR;
  pred_y = normal_rng(pred_mu, sigma);

  for(i in 1:N){
    log_lik[i] = normal_lpdf(y[i] | pred_mu[i], sigma);
  }
}
"
```

```{r}
q1_with_sey = list(N = nrow(full_data),
                y = full_data$log_gdp_std,
                A = full_data$cont_africa,
                R = full_data$rugged_std,
                AR = full_data$AR)

q1_no_sey = list(N = nrow(data_without_sey),
                y = data_without_sey$log_gdp_std,
                A = data_without_sey$cont_africa,
                R = data_without_sey$rugged_std,
                AR = data_without_sey$AR)

q1_sey_fit1 = stan(model_code = q1_m1,
               data = q1_with_sey,
               iter = 2000,
               chains = 2)

q1_no_sey_fit2 = stan(model_code = q1_m1,
               data = q1_no_sey,
               iter = 2000,
               chains = 2)
```
### result
$\beta_{AR}$ represents the effect of ruggedness depends upon continent, and both $\beta_{AR}$ in the two models(one fits to the full data, the other one fits to the data without country Seychelles.) have 89% HPDI that contain 0. Therefore, it seems there exist no statistically significant effect of ruggedness depending upon continent.  

Increasing 1 unit of ruggedness in any african country, we expect 0.0442 unit increase in y, which is log(gdp)/mean(log_gdp). Likewise, increasing 1 unit of ruggedness in african country except Seychelles, we expect 0.0322 unit increase in y, which is log(gdp)/mean(log_gdp).
```{r}
q1_a_result1 = as.data.frame(q1_sey_fit1)
q1_a_result2 = as.data.frame(q1_no_sey_fit2)

beta_ar1 = data.frame(
  mean = q1_a_result1 %>% dplyr::select(contains("beta_ar")) %>% apply(.,2,mean),
  L_HPDI = q1_a_result1 %>% dplyr::select(contains("beta_ar")) %>% apply(., 2, HPDI) %>% .[1,],
  R_HPDI = q1_a_result1 %>% dplyr::select(contains("beta_ar")) %>% apply(., 2, HPDI) %>% .[2,]
)

beta_ar2 = data.frame(
  mean = q1_a_result2 %>% dplyr::select(contains("beta_ar")) %>% apply(.,2,mean),
  L_HPDI = q1_a_result2 %>% dplyr::select(contains("beta_ar")) %>% apply(., 2, HPDI) %>% .[1,],
  R_HPDI = q1_a_result2 %>% dplyr::select(contains("beta_ar")) %>% apply(., 2, HPDI) %>% .[2,]
)

print(beta_ar1)
print(beta_ar2)
```
 

## Part B
plot the predictions
I consider that I get the same conclusion as the above Part A.
```{r}
pred_mu = q1_a_result1 %>% dplyr::select(contains("pred_mu"))
pred_y = q1_a_result1 %>% dplyr::select(contains("pred_y"))

result = data.frame(
  pred_mu = pred_mu %>% apply(., 2, mean),
  CI_lower = pred_mu %>% apply(., 2, HPDI) %>% .[1,],
  CI_upper = pred_mu %>% apply(., 2, HPDI) %>% .[2,],
  PI_lower = pred_y %>% apply(., 2, HPDI) %>% .[1,],
  PI_upper = pred_y %>% apply(., 2, HPDI) %>% .[2,],
  gdp = full_data$log_gdp_std,
  rug = full_data$rugged_std,
  africa = full_data$cont_africa
)

picture_b= result %>% 
  ggplot(aes(rug, gdp, fill = ifelse(africa==1, "africa", "else"))) +
  geom_point(shape=21, stroke=0) + # stroke defines the thickness of the border of the point
  geom_line(aes(rug, pred_mu)) +
  geom_ribbon(aes(x=rug, ymin=CI_lower, ymax=CI_upper), alpha=.6) +
  labs(fill="Continent")

picture_b
picture_b + facet_wrap(~africa)
```

## Part C
### Model1
```{r}
q1_c_m1 <- "
data{
  int N;
  vector[N] y;
  vector[N] R;
}

parameters{
  real alpha;
  real beta_r;
  real sigma;
}

model{
  vector[N] mu = alpha + beta_r*R;
  y ~ normal(mu, sigma);

  alpha ~ normal(1, 0.1);
  beta_r ~ normal(0, 0.3);
  sigma ~ exponential(1);
}

generated quantities{
  vector[N] pred_mu;
  vector[N] log_lik;

  pred_mu = alpha + beta_r*R;

  for(i in 1:N){
    log_lik[i] = normal_lpdf(y[i] | pred_mu[i], sigma);
  }
}
"
```

```{r}
q1_c_data1 = list(N = nrow(data_without_sey),
                  y = log(data_without_sey$rgdppc_2000),
                  R = data_without_sey$rugged)

q1_c_fit1 = stan(model_code = q1_c_m1,
                 data = q1_c_data1,
                 iter = 4000,
                 chains = 1)
```


### Model2
```{r}

q1_c_m2 <- "
data{
  int N;
  vector[N] y;
  vector[N] A;
  vector[N] R;
}

parameters{
  real alpha;
  real beta_a;
  real beta_r;
  real sigma;
}

model{
  vector[N] mu = alpha + beta_a*A + beta_r*R;
  y ~ normal(mu, sigma);

  alpha ~ normal(1, 0.1);
  beta_a ~ normal(0, 0.3);
  beta_r ~ normal(0, 0.3);
  sigma ~ exponential(1);
}

generated quantities{
  vector[N] pred_mu;
  vector[N] log_lik;

  pred_mu = alpha + beta_a*A + beta_r*R;

  for(i in 1:N){
    log_lik[i] = normal_lpdf(y[i] | pred_mu[i], sigma);
  }
}
"
```

```{r}
q1_c_data2 = list(N = nrow(data_without_sey),
                  y = log(data_without_sey$rgdppc_2000),
                  A = data_without_sey$cont_africa,
                  R = data_without_sey$rugged)

q1_c_fit2 = stan(model_code = q1_c_m2,
                 data = q1_c_data2,
                 iter = 4000,
                 chains = 1)
```


### Model3
```{r}

q1_c_m3 ="
data{
  int N;
  vector[N] y;
  vector[N] A;
  vector[N] R;
  vector[N] AR;
}

parameters{
  real alpha;
  real beta_a;
  real beta_r;
  real beta_ar;
  real sigma;
}

model{
  vector[N] mu = alpha + beta_a*A + beta_r*R + beta_ar*AR;
  y ~ normal(mu, sigma);

  alpha ~ normal(1, 0.1);
  beta_a ~ normal(0, 0.3);
  beta_r ~ normal(0, 0.3);
  beta_ar ~ normal(0, 0.3);
  sigma ~ exponential(1);
}

generated quantities{
  vector[N] pred_mu;
  vector[N] log_lik;

  pred_mu = alpha + beta_a*A + beta_r*R + beta_ar*AR;

  for(i in 1:N){
    log_lik[i] = normal_lpdf(y[i] | pred_mu[i], sigma);
  }
}
"
```

```{r}
q1_c_data3 = list(N = nrow(data_without_sey),
                  y = log(data_without_sey$rgdppc_2000),
                  A = data_without_sey$cont_africa,
                  R = data_without_sey$rugged,
                  AR = data_without_sey$AR)

q1_c_fit3 = stan(model_code = q1_c_m3,
                 data = q1_c_data3,
                 iter = 4000,
                 chains = 1)
```

### WAIC
Model 2 fits the best, since it has the lowest WAIC (slightly smaller than model 3), and the result also tells us that adding interaction doesn't benefit model to explain the data.  
```{r}
compare(q1_c_fit1, q1_c_fit2, q1_c_fit3)
```


```{r}
theta[4,1]
```

## Part D
```{r}
set.seed(1)
x = q1_c_data3$y  ## simulated data 
n.rep = 10000
n.accept = 0

A = q1_c_data3$A
R = q1_c_data3$R
AR = q1_c_data3$AR

theta <- matrix(0, 5, n.rep) ## storage for sim. values 
ll0 <- sum(dnorm(
              x,
              mean = theta[1,1] + theta[2,1]*A + theta[3,1]*R + theta[4,1]*AR,
              sd=exp(theta[5,1]),
              log=TRUE)) 

for (i in 2:n.rep) { ## The MH loop
  theta[,i] <- theta[,i-1] + rt(5,df=100)*0.05 ## proposal 
  ll1 <- sum(dnorm(
              x,
              mean = theta[1,i] + theta[2,i]*A + theta[3,i]*R + theta[4,i]*AR,
              sd=exp(theta[5,i]),
              log=TRUE)) 
  if (exp(ll1-ll0)>runif(1)) { ## MH accept/reject
    ll0 = ll1
    n.accept = n.accept+1 ## accept 
  }
  else theta[,i] = theta[,i-1] ## reject
}

```

```{r}
layout(matrix(1:8, 2, 4))

plot(1:n.rep,
     theta[1,], 
     type="l",
     xlab="iteration",
     ylab="alpha")

hist(theta[1,-(1:1000)],main="", xlab="alpha")

plot(1:n.rep,
     theta[2,],
     type="l",
     xlab="iteration",
     ylab="beta_A")

hist(theta[2,-(1:1000)],main="", xlab="beta_A")

plot(1:n.rep,
     theta[3,], 
     type="l",
     xlab="iteration",
     ylab="beta_R")

hist(theta[3,-(1:1000)],main="", xlab="beta_R")

plot(1:n.rep,
     theta[4,],
     type="l",
     xlab="iteration",
     ylab="beta_AR")

hist(theta[4,-(1:1000)],main="", xlab="beta_AR")

```

I drop the first 1000 sigma, because they are unstable

```{r}
layout(matrix(1:2, 1, 2))

plot(1001:n.rep,
     exp(theta[5,-(1:1000)]),
     type="l",
     xlab="iteration",
     ylab=expression(sigma))

hist(exp(theta[5,-(1:1000)]),main="", xlab=expression(sigma))

```

