---
title: "MA_Hw_04_02"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---
# Import library

```{r}
suppressMessages(library(plyr))
suppressMessages(library(rethinking))
suppressMessages(library(rstan))
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
```

# Assigned Part
required

```{r}
lr.model="
data{
    int <lower=1> N;
}
parameters{
    real a;
    real b;
}
model{
    a ~ normal(0,1);
    b ~ cauchy(0,1);
}
"
```

```{r}
lr.fit= stan(
            model_code = lr.model,
            data = list(N=1),
            chain = 2,
            cores = 2,
            iter = 2000)
lr.post = as.data.frame(lr.fit)
```

# Conclusion
We sampling two distribution, normal distribution and Cauchy distribution.  
With the summary and the following pictures, you can see that normal distribution and Cauchy distribuion are a little similar to each other. However, you should note that normal distribution are more centerlised, and also Cauchy distribution's tails relatively decay much more slowly.

```{r}
summary(lr.post)

q2_ab = ggplot()+
  geom_histogram(aes(lr.post$a, color = "normal"), bins = 1000, alpha=1)+
  geom_histogram(aes(lr.post$b, color = "Cauchy"), bins = 1000, alpha=0.01)

q2_ab
```

