---
title: "MA_Hw_06_02"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---

## Packages
```{r}
suppressMessages(library(plyr))
suppressMessages(library(rethinking))
suppressMessages(library(rstan))
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(skimr)) # for n_unique func
```

## Data-processing
### Read data
```{r}
data(Oxboys)
Oxboys
```

### Select data
I select the data which might be used first
```{r}
train = Oxboys %>%
  dplyr::select("Subject", "age", "height")
train
```

## PART A
### Centered Model
Centered varying model
```{r}
q2_model_A = "
data{
	int N;
	int N_boys;
	
	int boy_id[N];
	real age[N];
	real height[N];
}
parameters{
	
	//parameters
	real alpha[N_boys];
	real beta[N_boys];
	real<lower=0> sigma_height;

	//hyperparameters
	real hyper_alpha;
	real hyper_beta;
	vector<lower=0>[2] sigma_par; // [sigma_alpha, sigma_beta]
	corr_matrix[2] R;

}
transformed parameters{

	//variable declarations
	vector[2] y[N_boys];	  // y vector, N_boys x 2 matrix
  vector[2] Mu;					  // Mu vector in MVN
	matrix[2, 2] Cov;			// Cov matrix in MVN
	vector[N] mu_height;			// mu_height

  // y vector
	for (j in 1:N_boys) {
		y[j, 1] = alpha[j];
		y[j, 2] = beta[j];

	  // Mu vector in MVN
	  Mu[1] = hyper_alpha;
	  Mu[2] = hyper_beta;
	
	  // Cov matrix in MVN
	  Cov = quad_form_diag(R, sigma_par); 
	  // i.e. diag_matrix(sigma_par)  * R * diag_matrix(sigma_par)
	
	}

	// mu_height
	for (i in 1:N){
		mu_height[i] = alpha[boy_id[i]] + beta[boy_id[i]] * age[i];
	}

}
model{
	
	//model
	height ~ normal(mu_height, sigma_height);

	//prior
	sigma_height ~ exponential(1);
	y ~ multi_normal(Mu, Cov);

	//hyper prior
	sigma_par ~ exponential(1);
	hyper_alpha ~ normal(0, 2);
	hyper_beta ~ normal(0, 0.5);
	R ~ lkj_corr(2);
	
}
generated quantities {
	vector[N] log_lik;
	for (i in 1:N){
		log_lik[i] = normal_lpdf(height[i] | mu_height[i], sigma_height);
	}
}
"
```


#### Data and Initialise
```{r}
q2_data_A = list(
  N = nrow(train),
	N_boys = length(unique(train$Subject)),
	
	boy_id = train$Subject,
	age = train$age,
  height = train$height
)
```
#### Fit
```{r}
q2_fit_A = stan(
                model_code = q2_model_A, 
                data = q2_data_A, 
                cores = 4, chains = 2, 
                warmup = 2000, iter = 8000,
                control = list(adapt_delta=.99)
                )

```

## PART B
### Un-Centered Model
To decompose the Cov, I use the L_Rho and Z_alpha to replace the original R and diag_matrix(sigma_par)
```{r}
q2_model_B = "
data {
	int N;
	int N_boys;
	
	int boy_id[N];
	real age[N];
	real height[N];
}

parameters {
	//hyperparameters
	cholesky_factor_corr[2] L_Rho_alpha;
	vector<lower=0>[2] sigma_alpha;

	//parameters
	matrix[2, N_boys] z_alpha;
	real<lower=0> sigma_height;
}

transformed parameters{
  vector[N] mu_height;
	matrix[N_boys, 2] alpha;

	// calculate alpha
	alpha = (diag_pre_multiply(sigma_alpha, L_Rho_alpha) * z_alpha)'; 
	// we transpose the result to make it a N_boys x 2 matrix

	// calculate height **you have to put this part after the parts calculating alpha/beta**
	for (i in 1:N){
		mu_height[i] = alpha[boy_id[i], 1] + alpha[boy_id[i], 2] * age[i];
	}
	
}
model{
  //model
	height ~ normal(mu_height, sigma_height);

	//prior
	sigma_height ~ exponential(1);
	to_vector(z_alpha) ~ normal(0, 1);

	//hyperprior
	L_Rho_alpha ~ lkj_corr_cholesky(2);
	sigma_alpha ~ exponential(1);

}
generated quantities{
	vector[N] log_lik;
	
	for (i in 1:N){
		log_lik[i] = normal_lpdf(height[i] | mu_height[i], sigma_height);
	}
}
"
```

#### Data and Initialise
```{r}
q2_data_B = list(
  N = nrow(train),
	N_boys = length(unique(train$Subject)),
	
	boy_id = train$Subject,
	age = train$age,
  height = train$height
)
```
#### Fit
```{r}
q2_fit_B = stan(
    model_code = q2_model_B, 
    data = q2_data_B, 
    cores = 4, chains = 2, 
    warmup = 2000, iter = 16000,
    control = list(adapt_delta=.99, max_treedepth = 25)
    )

```

## PART C
### Check the posteriors
we can see that all the R hat of paras of both models are 1, which means they are convergent
```{r}
options(max.print=1000000)
print(q2_fit_A, pars = c( "sigma_par"))
print(q2_fit_B, pars = c( "sigma_alpha"))
```

### Variation to the heights
By seeing the sigmas of two models,  
we can see that the <span style="color: blue;">`first one`</span> in sigma is <span style="color: blue;">`larger`</span> than the second one in both models.  
The first one in both models represent the <span style="color: blue;">`intercept`</span>,  
and hence, we could know <span style="color: blue;">`intercept`</span> contributes more variation to the heights.